---
title: "Step1 - Complete Data"
output: html_document
---

## Goal of Script
1. Download All Data
# per object
2. Run checkDataObject(dat$originalData$dataset, "Dataset") per object and check
    a. The pheno is there
    b. The expr is there
    c. The keys are there
    d. The colnames(expr) == rownames(pheno) and keys == rownames(expr)
    e. Check normalization and naming of keys

```{r load_libraries}
library(MetaIntegrator)
library(tidyr)
library(stringr)
library(tximport)
library(dplyr)
library(tximportData)
library(DESeq2)
library(tidyverse)
library(data.table)
library(HGNChelper)

set.seed(4)
```

```{r setup_paths}
path = paste0(here::here(),"/") ## should be the github project folder, otherwise adjust accordingly

datPath = paste0(path, "data/")
datPath_raw = paste0(datPath, "raw_data/")
datPath_p = paste0(datPath, "processed_data/")
dataPath_input = paste0(datPath, "raw_input/")

## create these new directories
dir.create(file.path(datPath))
dir.create(file.path(datPath_raw))
dir.create(file.path(datPath_p))
dir.create(file.path(dataPath_input))

## download method
download_method = "curl" ## wget, libcurl, etc.
```


## Download and save data
- added extras as placeholder for Sedar and PRJNAs
```{r}
# use R to directly download files from Baize and GSE184949_GPL29319

# use getGeo to download data
dat <- getGEOData(c("GSE41752", "GSE184949")) 

# save original data
saveRDS(dat, file = paste0(datPath_raw, "01_original_download_data.rds"))
saveRDS(dat$originalData$GSE41752, file = paste0(datPath_raw, "GSE41752","_originalData.rds"))
saveRDS(dat$originalData$GSE184949_GPL29319, file = paste0(datPath_raw, "GSE184949_GPL29319","_originalData.rds"))
saveRDS(dat$originalData$GSE184949_GPL29759, file = paste0(datPath_raw, "GSE184949_GPL29759","_originalData.rds"))
```

## Check each data object/ fill in"

# Key fix from here: https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?view=data&acc=GPL14569&id=7598&db=GeoDb_blob92 
```{r}
dat = readRDS(file = paste0(datPath_raw, "01_original_download_data.rds"))
```

### Check 1. GSE41752-LASV_1
```{r GSE41752-LASV_1}
dat$originalData$GSE41752$name = "LASV_1"

# check for dates & normalization
dobj = dat$originalData$GSE41752
keys = dobj$keys
genes = unname(unlist(keys))
genes[grepl("^[[:digit:]]+",genes)]
boxplot(dobj$expr[,1:10], xlab = dobj$name)

checkDataObject(dat$originalData$GSE41752, "Dataset") 
```

### Check 2. GSE184949-SARS_5
```{r GSE184949}
in_file = paste0(dataPath_input, "GSE184949/")
dir.create(file.path(in_file))
# ##download files
download.file("https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE184949&format=file", destfile = paste0(in_file,"counts_matrices.tar"), method = download_method)
untar(paste0(in_file,"counts_matrices.tar"), exdir = paste0(in_file,"counts_matrices"))

## add files into geo object and process
dat$originalData$GSE184949_GPL29319$name = "SARS_5"
checkDataObject(dat$originalData$GSE184949_GPL29319, "Dataset") 

pheno = dat$originalData$GSE184949_GPL29319$pheno
pheno$animal = substr(pheno$title, 1, 4)

dir1 = paste0(in_file, "counts_matrices/")
samples <- list.files(dir1)
samples = samples[grepl("GSM",samples)]
sample_table = data.frame(samples= samples)
sample_table$sample_id = substr(sample_table$samples, 1, 10)
sample_table = sample_table[(sample_table$sample_id %in% rownames(pheno)),]

file1 = read.table(paste0(dir1, samples[1]), col.names=c("gene_id",samples[1]))
file2 = read.table(paste0(dir1, samples[2]), col.names=c("gene_id",samples[2]))
out.file = merge (file1, file2, by=c("gene_id"))

for(i in 3:length(samples)){
  file = read.table(paste0(dir1, samples[i]),col.names=c("gene_id",samples[i]))
  out.file <- merge(out.file, file, by=c("gene_id"))
}

rownames(sample_table) = sample_table$sample_id
sample_table = merge(sample_table, pheno, by = 0)

sampleTable <- data.frame(sampleName = sample_table$sample_id,
                      fileName = sample_table$samples,
                      condition = sample_table$animal)


ddsHTSeq <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
                                   directory = dir1,
                                   design= ~ condition)

out_list = rownames(ddsHTSeq@assays@data$counts)[grepl("N_", rownames(ddsHTSeq@assays@data$counts))]

vsd <- vst(ddsHTSeq, blind=FALSE)
normalizedGeneCounts = assay(vsd)
normalizedGeneCounts = normalizedGeneCounts[!rownames(normalizedGeneCounts) %in% out_list,]
table(duplicated(rownames(normalizedGeneCounts)))

dat$originalData$GSE184949_GPL29319$expr = normalizedGeneCounts
dat$originalData$GSE184949_GPL29319$keys = rownames(normalizedGeneCounts)
names(dat$originalData$GSE184949_GPL29319$keys) = rownames(normalizedGeneCounts)

# check for dates & count normalization
dobj = dat$originalData$GSE184949_GPL29319
keys = dobj$keys
genes = unname(unlist(keys))
genes[grepl("^[[:digit:]]+",genes)]
boxplot(dobj$expr[,1:10], xlab = dobj$name)

checkDataObject(dat$originalData$GSE184949_GPL29319, "Dataset") 
```

### Check 3. GSE184949-SARS_6
```{r GSE184949}
dat$originalData$GSE184949_GPL29759$name = "SARS_5"
checkDataObject(dat$originalData$GSE184949_GPL29759, "Dataset") 

pheno = dat$originalData$GSE184949_GPL29759$pheno
pheno$animal = substr(pheno$title, 1, 4)

## add files into geo object and process
in_file = paste0(dataPath_input, "GSE184949/")
dir1 = paste0(in_file, "counts_matrices/")
samples <- list.files(dir1)
samples = samples[grepl("GSM",samples)]
sample_table = data.frame(samples= samples)
sample_table$sample_id = substr(sample_table$samples, 1, 10)
sample_table = sample_table[(sample_table$sample_id %in% rownames(pheno)),]

file1 = read.table(paste0(dir1, samples[1]), col.names=c("gene_id",samples[1]))
file2 = read.table(paste0(dir1, samples[2]), col.names=c("gene_id",samples[2]))
out.file = merge (file1, file2, by=c("gene_id"))

for(i in 3:length(samples)){
  file = read.table(paste0(dir1, samples[i]),col.names=c("gene_id",samples[i]))
  out.file <- merge(out.file, file, by=c("gene_id"))
}

rownames(sample_table) = sample_table$sample_id
sample_table = merge(sample_table, pheno, by = 0)

sampleTable <- data.frame(sampleName = sample_table$sample_id,
                      fileName = sample_table$samples,
                      condition = sample_table$animal)


ddsHTSeq <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
                                   directory = dir1,
                                   design= ~ condition)

out_list = rownames(ddsHTSeq@assays@data$counts)[grepl("N_", rownames(ddsHTSeq@assays@data$counts))]

vsd <- vst(ddsHTSeq, blind=FALSE)
normalizedGeneCounts = assay(vsd)
normalizedGeneCounts = normalizedGeneCounts[!rownames(normalizedGeneCounts) %in% out_list,]
table(duplicated(rownames(normalizedGeneCounts)))

dat$originalData$GSE184949_GPL29759$expr = normalizedGeneCounts
dat$originalData$GSE184949_GPL29759$keys = rownames(normalizedGeneCounts)
names(dat$originalData$GSE184949_GPL29759$keys) = rownames(normalizedGeneCounts)

# check for dates & count normalization
dobj = dat$originalData$GSE184949_GPL29759
keys = dobj$keys
genes = unname(unlist(keys))
genes[grepl("^[[:digit:]]+",genes)]
boxplot(dobj$expr[,1:10], xlab = dobj$name)

checkDataObject(dat$originalData$GSE184949_GPL29759, "Dataset") 
```

### Check 4. Baize-MACV_1
```{r Baize-MACV_1}
in_file = paste0(dataPath_input, "Baize/")
dir.create(file.path(in_file))

## download data
download.file('https://zenodo.org/records/7229439/files/counts_post.txt?download=1', destfile = paste0(in_file,"counts_post.txt"), method = download_method)
download.file('https://zenodo.org/records/7229439/files/target_post.txt?download=1', destfile = paste0(in_file,"post-infection-timepoints.csv"), method = download_method)

## add files into geo object and process
dat$originalData$Baize = dat$originalData$GSE99463
dat$originalData$Baize$name = "MACV_1"
obj = dat$originalData$Baize

metadata = read.table(paste0(in_file,"post-infection-timepoints.csv"), sep = " ", header = T)
rownames(metadata) = metadata$samples
metadata = separate(metadata, col = "samples",sep = "_", into = c("animal","day"), remove = FALSE)
metadata$day = gsub("J","", metadata$day)


require(biomaRt)
ensembl <- useMart('ensembl', dataset = 'mfascicularis_gene_ensembl')
annot <- getBM(
  attributes = c(
    'external_gene_name',
    'ensembl_gene_id',
    'gene_biotype',
    'external_synonym'),
  mart = ensembl)

saveRDS(annot, paste0(in_file,"annotation_cm.rds"))
annot = readRDS(paste0(in_file,"annotation_cm.rds"))

expr_dat = read.table(paste0(in_file,"counts_post.txt"), row.names = 1)
dds <- DESeqDataSetFromMatrix(countData = expr_dat,
                              colData = metadata,
                              design = ~ animal)

vsd <- vst(dds, blind=FALSE)
normalizedGeneCounts = assay(vsd)

# create key
annot_order = annot
annot_order$external_synonym = NULL
annot_order$gene_biotype = NULL
annot_order = unique(annot_order)
table(duplicated(annot_order$ensembl_gene_id))

df = rownames(normalizedGeneCounts) %>% as.data.frame()
colnames(df) = "ensembl_gene_id"
key = merge(df, annot_order, by = "ensembl_gene_id", all.x = TRUE, all.y = FALSE)

key = key[match(rownames(normalizedGeneCounts), key$ensembl_gene_id),]
setequal(rownames(normalizedGeneCounts), key$ensembl_gene_id)
rownames(key) = key$ensembl_gene_id
key$ensembl_gene_id = NULL
###########
obj$pheno = metadata
obj$expr = normalizedGeneCounts
obj$keys = key$external_gene_name
names(obj$keys) = rownames(key)
obj$name = "Baize"
obj$platform = "Illumina"
obj$formattedName = "Baize"
obj$class = rep(0,nrow(metadata))
names(obj$class) = (rownames(metadata))

checkDataObject(obj, "Dataset") 

# check for dates & count normalization
dobj = obj
keys = dobj$keys
genes = unname(unlist(keys))
genes[grepl("^[[:digit:]]+",genes)]
boxplot(dobj$expr[,1:10], xlab = dobj$name)

dat$originalData$Baize = obj
```

```{r order_object}
dat_ordered = list()
dat_ordered$originalData = list()
dat_ordered$originalData$Baize = dat$originalData$Baize
dat_ordered$originalData$GSE41752 = dat$originalData$GSE41752
dat_ordered$originalData$GSE184949_GPL29319 = dat$originalData$GSE184949_GPL29319
dat_ordered$originalData$GSE184949_GPL29759 = dat$originalData$GSE184949_GPL29759

length(names(dat_ordered$originalData)) == sum(names(dat_ordered$originalData) %in% names(dat$originalData))
```

```{r save_object}
checkDataObject(dat, "Meta", "Pre-Analysis")
saveRDS(dat, file = paste0(datPath_p, "1a_original_completed_data.rds"))
checkDataObject(dat_ordered, "Meta", "Pre-Analysis")
saveRDS(dat_ordered, file = paste0(datPath_p, "1b_original_completed_ordered_data.rds"))
```