---
title: "01_3A"
output: html_document
---
---
title: "Generating Figure 3A"
output: html_document
---
Thresholding DEGs: meta_fdr < .05, effect size >= .1, present in at least 2 datasets (Flavi broken into 4) -- purpose to look for BROAD DEGs for discovery
1. Generated 14a_timecats_list.rds - list of objects broken up by time_category
2. Generated 15a_viruslist_t1, 15a_viruslist_t2 etc. - list of time category objects broken by virus
3. virus..._timesplit_dat.rds = list of timepoint splits per virus (reverse of 2)
4. Generated 3A_deg_count_table.rds, DEGs by virus x time category
5. Generated 3A figure

## script to plot each object by MVS score/ day and make table of mvs counts
```{r}
library(MetaIntegrator)
library(tidyr)
library(stringr)
library(dplyr)
# library(DESeq2)
library(tidyverse)
library(data.table)
library(ggpubr)
library(NatParksPalettes)

set.seed(4)
```

```{r setup_paths}
path = ""
datPath = paste0(path, "data/processed_data/")
datTable = paste0(path, "data/data_tables/")
source(paste0(path,"scripts/load_mvs_genes.R"))
source(paste0(path,"scripts/functions_step45.R"))
source(paste0(path,"scripts/subset_dataobjects.R"))
fig_path = paste0(path,"figures/")
fig3_path = paste0(fig_path, "Figure3/")
```

## starting input
```{r}
dat = readRDS(paste0(datPath, "5a_scores_added.rds"))
```

```{r}
## pick the earliest day in the time category
subset_data_by_earliest_time = function(object) {
  data_out = list()
  object = fix_time_cat(object)
  for (i in 1:length(names(object$originalData))) {
    dataset = object$originalData[[i]]
    pheno = dataset$pheno
    name = unname(unlist(unique(pheno[colname])))
    df_one = df[df$obj_name == name,]
    df_one$ani_day = paste0(df_one$animal,"_",df_one$day)
    pheno$ani_day = paste0(pheno$animal,"_",pheno$day)

    ## subset pheno by animal+time
    pheno_sub = pheno[pheno$ani_day %in% df_one$ani_day,]
    print(paste0(name,"_",nrow(pheno),"_",nrow(pheno_sub)))
    ## subset by dataset by pheno
    dataset <- subsetOriginalData(dataset, 
                                  keepMe= c(rownames(pheno_sub)))
    message(checkDataObject(dataset, "Dataset"))
    object$originalData[[i]] = dataset
    }
  return(object)
}
```

## subset data by time_category
```{r}
dat_t1 = subset_by_timecat(dat, "t1")
dat_t2 = subset_by_timecat(dat, "t2")
dat_t3 = subset_by_timecat(dat, "t3")
dat_t4 = subset_by_timecat(dat, "t4")
dat_t5 = subset_by_timecat(dat, "t5")

timecat_list = list(dat_t1, dat_t2, dat_t3, dat_t4, dat_t5)
names(timecat_list) = c("t1", "t2", "t3", "t4", "t5")
saveRDS(timecat_list, file = paste0(datPath, "14a_timecats_list.rds"))
```


```{r}
timecat_list = readRDS(paste0(datPath, "14a_timecats_list.rds"))
dat_t1 = timecat_list$t1 %>% remove_empties(time = "t1")
dat_t2 = timecat_list$t2 %>% remove_empties(time = "t2")
dat_t3 = timecat_list$t3 %>% remove_empties(time = "t3")
dat_t4 = timecat_list$t4 %>% remove_empties(time = "t4")
dat_t5 = timecat_list$t5 %>% remove_empties(time = "t5")

arena_dat = reformat(dat_t1, arena,  reMetanalaysis = TRUE)
arena_dat  = fix_formattedname(arena_dat)
corona_dat = reformat(dat_t1, corona,  reMetanalaysis = TRUE)
corona_dat  = fix_formattedname(corona_dat)
filo_dat = reformat(dat_t1, filo,  reMetanalaysis = TRUE)
filo_dat  = fix_formattedname(filo_dat)
flavi_dat = reformat(dat_t1, flavi,  reMetanalaysis = TRUE)
flavi_dat  = fix_formattedname(flavi_dat)
ortho_dat = reformat(dat_t1, ortho,  reMetanalaysis = TRUE)
ortho_dat  = fix_formattedname(ortho_dat)

virus_dat_list_t1 = list(arena_dat, corona_dat, filo_dat, flavi_dat, ortho_dat)
names(virus_dat_list_t1) = c("arena", "corona", "filo", "flavi", "ortho")
saveRDS(virus_dat_list_t1, file = paste0(datPath, "15a_viruslist_t1.rds"))



arena_dat = reformat(dat_t2, arena,  reMetanalaysis = TRUE)
arena_dat  = fix_formattedname(arena_dat)
corona_dat = reformat(dat_t2, corona,  reMetanalaysis = TRUE)
corona_dat  = fix_formattedname(corona_dat)
filo_dat = reformat(dat_t2, filo,  reMetanalaysis = TRUE)
filo_dat  = fix_formattedname(filo_dat)
flavi_dat = reformat(dat_t2, flavi,  reMetanalaysis = TRUE)
flavi_dat  = fix_formattedname(flavi_dat)
ortho_dat = reformat(dat_t2, ortho,  reMetanalaysis = TRUE)
ortho_dat  = fix_formattedname(ortho_dat)

virus_dat_list_t2 = list(arena_dat, corona_dat, filo_dat, flavi_dat, ortho_dat)
names(virus_dat_list_t2) = c("arena", "corona", "filo", "flavi", "ortho")
saveRDS(virus_dat_list_t2, file = paste0(datPath, "15a_viruslist_t2.rds"))


arena_dat = reformat(dat_t3, arena,  reMetanalaysis = TRUE)
arena_dat  = fix_formattedname(arena_dat)
corona_dat = reformat(dat_t3, corona,  reMetanalaysis = TRUE)
corona_dat  = fix_formattedname(corona_dat)
filo_dat = reformat(dat_t3, filo,  reMetanalaysis = TRUE)
filo_dat  = fix_formattedname(filo_dat)
flavi_dat = reformat(dat_t3, flavi,  reMetanalaysis = TRUE)
flavi_dat  = fix_formattedname(flavi_dat)
ortho_dat = reformat(dat_t3, ortho,  reMetanalaysis = TRUE)
ortho_dat  = fix_formattedname(ortho_dat)

virus_dat_list_t3 = list(arena_dat, corona_dat, filo_dat, flavi_dat, ortho_dat)
names(virus_dat_list_t3) = c("arena", "corona", "filo", "flavi", "ortho")
saveRDS(virus_dat_list_t3, file = paste0(datPath, "15a_viruslist_t3.rds"))

arena_dat = reformat(dat_t4, arena,  reMetanalaysis = TRUE)
arena_dat  = fix_formattedname(arena_dat)
corona_dat = reformat(dat_t4, corona,  reMetanalaysis = TRUE)
corona_dat  = fix_formattedname(corona_dat)
filo_dat = reformat(dat_t4, filo,  reMetanalaysis = TRUE)
filo_dat  = fix_formattedname(filo_dat)
# flavi_dat = reformat(dat_t4, flavi,  reMetanalaysis = TRUE)
# flavi_dat  = fix_formattedname(flavi_dat)
# ortho_dat = reformat(dat_t4, ortho,  reMetanalaysis = TRUE)
# ortho_dat  = fix_formattedname(ortho_dat)

virus_dat_list_t4 = list(arena_dat, corona_dat, filo_dat)
names(virus_dat_list_t4) = c("arena", "corona", "filo")
saveRDS(virus_dat_list_t4, file = paste0(datPath, "15a_viruslist_t4.rds"))


arena_dat = reformat(dat_t5, arena,  reMetanalaysis = TRUE)
arena_dat  = fix_formattedname(arena_dat)
corona_dat = reformat(dat_t5, corona,  reMetanalaysis = TRUE)
corona_dat  = fix_formattedname(corona_dat)
filo_dat = reformat(dat_t5, filo,  reMetanalaysis = TRUE)
flavi_dat = reformat(dat_t5, flavi,  reMetanalaysis = TRUE)
flavi_dat  = fix_formattedname(flavi_dat)
filo_dat  = fix_formattedname(filo_dat)
ortho_dat = reformat(dat_t5, ortho,  reMetanalaysis = TRUE)
ortho_dat  = fix_formattedname(ortho_dat)

virus_dat_list_t5 = list(arena_dat, corona_dat, filo_dat, flavi_dat, ortho_dat)
names(virus_dat_list_t5) = c("arena", "corona", "filo","flavi",  "ortho")
saveRDS(virus_dat_list_t5, file = paste0(datPath, "15a_viruslist_t5.rds"))
```

```{r}
virus_dat_list_t1 = readRDS(paste0(datPath, "15a_viruslist_t1.rds"))
virus_dat_list_t2 = readRDS(paste0(datPath, "15a_viruslist_t2.rds"))
virus_dat_list_t3 = readRDS(paste0(datPath, "15a_viruslist_t3.rds"))
virus_dat_list_t4 = readRDS(paste0(datPath, "15a_viruslist_t4.rds"))
virus_dat_list_t5 = readRDS(paste0(datPath, "15a_viruslist_t5.rds"))

arena_t1_dat = virus_dat_list_t1$arena
arena_t2_dat = virus_dat_list_t2$arena
arena_t3_dat = virus_dat_list_t3$arena
arena_t4_dat = virus_dat_list_t4$arena
arena_t5_dat = virus_dat_list_t5$arena

corona_t1_dat = virus_dat_list_t1$corona
corona_t2_dat = virus_dat_list_t2$corona
corona_t3_dat = virus_dat_list_t3$corona
corona_t4_dat = virus_dat_list_t4$corona
corona_t5_dat = virus_dat_list_t5$corona

filo_t1_dat = virus_dat_list_t1$filo
filo_t2_dat = virus_dat_list_t2$filo
filo_t3_dat = virus_dat_list_t3$filo
filo_t4_dat = virus_dat_list_t4$filo
filo_t5_dat = virus_dat_list_t5$filo

flavi_t1_dat = virus_dat_list_t1$flavi
flavi_t2_dat = virus_dat_list_t2$flavi
flavi_t3_dat = virus_dat_list_t3$flavi
flavi_t4_dat = virus_dat_list_t4$flavi
flavi_t5_dat = virus_dat_list_t5$flavi

ortho_t1_dat = virus_dat_list_t1$ortho
ortho_t2_dat = virus_dat_list_t2$ortho
ortho_t3_dat = virus_dat_list_t3$ortho
ortho_t4_dat = virus_dat_list_t4$ortho
ortho_t5_dat = virus_dat_list_t5$ortho
```



```{r}
padj_thresh = 0.05 ## threshold for fdr threshold for usable genes 
es_thresh = 0.1
min_studies = 2

## cleaning pooled results for deg count
threholding_deg_data = function(dat_obj, padj.threshold = padj_thresh, es.threshold = es_thresh, studies.threshold = min_studies, virus = "", time_cat = "") {
  pooled_results_df = dat_obj$metaAnalysis$pooledResults
  pooled_results_df$direction = ifelse(pooled_results_df$effectSize > 0, "pos", "neg")
  pooled_results_df$gene = rownames(pooled_results_df)
  pooled_results_df = pooled_results_df[pooled_results_df$effectSizeFDR < padj.threshold & abs(pooled_results_df$effectSize) >= es.threshold & pooled_results_df$numStudies >= studies.threshold,]
  pooled_results_df = as.data.frame(pooled_results_df)
  pooled_results_df$Virus = virus
  pooled_results_df$time_cat = time_cat
  print(table(pooled_results_df$direction))
  return(pooled_results_df)
}
```

## subsetting pooled results
```{r}
# arena
arena_results_t1 = threholding_deg_data(arena_t1_dat, virus = "Arenaviridae",time_cat = "T1")
arena_results_t2 = threholding_deg_data(arena_t2_dat, virus = "Arenaviridae",time_cat = "T2")
arena_results_t3 = threholding_deg_data(arena_t3_dat, virus = "Arenaviridae",time_cat = "T3")
arena_results_t4 = threholding_deg_data(arena_t4_dat, virus = "Arenaviridae",time_cat = "T4")
# arena_results_t5 = threholding_deg_data(arena_t5_dat, virus = "Arenaviridae",time_cat = "T5")

# corona
corona_results_t1 = threholding_deg_data(corona_t1_dat, virus = "Coronaviridae",time_cat = "T1")
corona_results_t2 = threholding_deg_data(corona_t2_dat, virus = "Coronaviridae",time_cat = "T2")
corona_results_t3 = threholding_deg_data(corona_t3_dat, virus = "Coronaviridae",time_cat = "T3")
# corona_results_t4 = threholding_deg_data(corona_t4_dat, virus = "Coronaviridae",time_cat = "T4")
corona_results_t5 = threholding_deg_data(corona_t5_dat, virus = "Coronaviridae",time_cat = "T5")

# filo
filo_results_t1 = threholding_deg_data(filo_t1_dat, virus = "Filoviridae",time_cat = "T1")
filo_results_t2 = threholding_deg_data(filo_t2_dat, virus = "Filoviridae",time_cat = "T2")
filo_results_t3 = threholding_deg_data(filo_t3_dat, virus = "Filoviridae",time_cat = "T3")
filo_results_t4 = threholding_deg_data(filo_t4_dat, virus = "Filoviridae",time_cat = "T4")
# filo_results_t5 = threholding_deg_data(filo_t5_dat, virus = "Filoviridae",time_cat = "T5")

# flavi
flavi_results_t1 = threholding_deg_data(flavi_t1_dat, virus = "Flavivirdae",time_cat = "T1")
flavi_results_t2 = threholding_deg_data(flavi_t2_dat, virus = "Flavivirdae",time_cat = "T2")
flavi_results_t3 = threholding_deg_data(flavi_t3_dat, virus = "Flavivirdae",time_cat = "T3")
# flavi_results_t4 = threholding_deg_data(flavi_t4_dat) # no data
# flavi_results_t5 = threholding_deg_data(flavi_t5_dat, virus = "Flavivirdae",time_cat = "T5")

# ortho
# ortho_results_t1 = threholding_deg_data(ortho_t1_dat, virus = "Orthomyxoviridae",time_cat = "T1")
ortho_results_t2 = threholding_deg_data(ortho_t2_dat, virus = "Orthomyxoviridae",time_cat = "T2")
ortho_results_t3 = threholding_deg_data(ortho_t3_dat, virus = "Orthomyxoviridae",time_cat = "T3")
# ortho_results_t4 = threholding_deg_data(ortho_t4_dat) # no data
# ortho_results_t5 = threholding_deg_data(ortho_t5_dat, virus = "Orthomyxoviridae",time_cat = "T5")


all_data = rbind(arena_results_t1,arena_results_t2,arena_results_t3,arena_results_t4,
                 corona_results_t1,corona_results_t2,corona_results_t3,corona_results_t5,
                 filo_results_t1,filo_results_t2,filo_results_t3,filo_results_t4,
                 flavi_results_t1,flavi_results_t2,flavi_results_t3,
                 ortho_results_t2,ortho_results_t3)
write.csv(all_data, file = paste0(datTable,"stable4_degs.csv"))
```



############### generate DEG count table

```{r}
cols = c("virus", "up", "down","total","dataset_num", "time_cat")
arena_df <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(arena_df) <- cols

virus = "Arenaviridae"
arena_list_pooled = list(arena_results_t1, arena_results_t2, arena_results_t3,arena_results_t4)#,arena_results_t5)
names(arena_list_pooled) = c("t1","t2","t3","t4")#,"t5")
arena_list_dat = list(arena_t1_dat, arena_t2_dat, arena_t3_dat,arena_t4_dat)#,arena_t5_dat)

for (i in c(1:length(arena_list_pooled))){
  item = arena_list_pooled[[i]]
  num_dfs = length(arena_list_dat[[i]]$originalData)
  time_name = names(arena_list_pooled)[i]
  row = c(virus,sum(item$direction == "pos"), sum(item$direction == "neg"), nrow(item), num_dfs, time_name)
  arena_df = rbind(arena_df, row)
  colnames(arena_df) = cols
}

corona_df <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(corona_df) <- cols

virus = "Coronaviridae"
corona_list_pooled = list(corona_results_t1, corona_results_t2, corona_results_t3,corona_results_t5)#corona_results_t4,
names(corona_list_pooled) = c("t1","t2","t3","t5")# v
corona_list_dat = list(corona_t1_dat, corona_t2_dat, corona_t3_dat,corona_t5_dat) #,corona_t4_dat

for (i in c(1:length(corona_list_pooled))){
  item = corona_list_pooled[[i]]
  num_dfs = length(corona_list_dat[[i]]$originalData)
  time_name = names(corona_list_pooled)[i]
  row = c(virus,sum(item$direction == "pos"), sum(item$direction == "neg"), nrow(item), num_dfs, time_name)
  corona_df = rbind(corona_df, row)
  colnames(corona_df) = cols
}


filo_df <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(filo_df) <- cols

virus = "Filoviridae"
filo_list_pooled = list(filo_results_t1, filo_results_t2, filo_results_t3,filo_results_t4) #,filo_results_t5)
names(filo_list_pooled) = c("t1","t2","t3","t4") #,"t5")
filo_list_dat = list(filo_t1_dat, filo_t2_dat, filo_t3_dat,filo_t4_dat)# ,filo_t5_dat)

for (i in c(1:length(filo_list_pooled))){
  item = filo_list_pooled[[i]]
  num_dfs = length(filo_list_dat[[i]]$originalData)
  time_name = names(filo_list_pooled)[i]
  row = c(virus,sum(item$direction == "pos"), sum(item$direction == "neg"), nrow(item), num_dfs, time_name)
  filo_df = rbind(filo_df, row)
  colnames(filo_df) = cols
}


flavi_df <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(flavi_df) <- cols

virus = "Flaviviridae"
flavi_list_pooled = list(flavi_results_t1, flavi_results_t2, flavi_results_t3)#,flavi_results_t5)
names(flavi_list_pooled) = c("t1","t2","t3")#,"t5")
flavi_list_dat = list(flavi_t1_dat, flavi_t2_dat, flavi_t3_dat)#,flavi_t5_dat)

for (i in c(1:length(flavi_list_pooled))){
  item = flavi_list_pooled[[i]]
  num_dfs = length(flavi_list_dat[[i]]$originalData)
  time_name = names(flavi_list_pooled)[i]
  row = c(virus,sum(item$direction == "pos"), sum(item$direction == "neg"), nrow(item), num_dfs, time_name)
  flavi_df = rbind(flavi_df, row)
  colnames(flavi_df) = cols
}



ortho_df <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(ortho_df) <- cols

virus = "Orthomyxoviridae"
ortho_list_pooled = list( ortho_results_t2, ortho_results_t3)#,ortho_results_t5) #ortho_results_t1,
names(ortho_list_pooled) = c("t2","t3")#,"t5") #"t1",
ortho_list_dat = list( ortho_t2_dat, ortho_t3_dat)#,ortho_t5_dat) #ortho_t1_dat,

for (i in c(1:length(ortho_list_pooled))){
  item = ortho_list_pooled[[i]]
  num_dfs = length(ortho_list_dat[[i]]$originalData)
  time_name = names(ortho_list_pooled)[i]
  row = c(virus,sum(item$direction == "pos"), sum(item$direction == "neg"), nrow(item), num_dfs, time_name)
  ortho_df = rbind(ortho_df, row)
  colnames(ortho_df) = cols
}




all_df = rbind(arena_df, corona_df, filo_df, flavi_df, ortho_df)
all_df$def_padj_thresh = padj_thresh
all_df$def_es_thresh = es_thresh
all_df$def_min_studies_thresh = min_studies

all_df$virus = factor(all_df$virus, levels = viral_families)
all_df$time_cat = toupper(all_df$time_cat)
all_df$time_cat = factor(all_df$time_cat, levels = toupper(c("t1","t2","t3","t4","t5")))
all_df$total = as.numeric(all_df$total)
all_df$up = as.numeric(all_df$up)
all_df$down = as.numeric(all_df$down)
saveRDS(all_df, paste0(datPath,"3A_deg_count_table.rds"))
```

```{r}
all_df = readRDS(paste0(datPath,"3A_deg_count_table.rds"))
colnames(all_df)[1] = "Virus"
all_df$total = as.numeric(all_df$total)
all_df_lim = all_df[all_df$total != 0,]
all_df_lim = all_df_lim[c("Virus","time_cat","up","down")]

all_df_long = melt(all_df_lim, id=c("Virus","time_cat"))
all_df_long$value = as.numeric(all_df_long$value)
```

# first level graph
```{r}
all_df_long_up = all_df_long[all_df_long$variable == "up",]
all_df_long_up["DEG Direction"] = all_df_long_up$variable
ggplot(all_df_long_up, aes(fill=`DEG Direction`, y=value, x=time_cat)) + 
    geom_bar(position="stack", stat="identity")  + theme_bw()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + facet_wrap(~Virus, scales = "free_y", nrow = 1) + ylab("Number of DEGs") + xlab("Time Category") + scale_fill_manual(values = c("up" = "#C0392B", "down"= "#1F618D"))
```

## final Fig3a
```{r  Fig3a}
plot = ggplot(all_df_long, aes(fill=`DEG Direction`, y=value, x=time_cat)) + 
    geom_bar(position="stack", stat="identity")  + theme_bw()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + facet_wrap(~Virus, scales = "free_y", nrow = 1) + ylab("Number of DEGs") + xlab("Time Category") + scale_fill_manual(values = c("up" = "#C0392B", "down"= "#1F618D"))+ theme(text = element_text(size = 16))
plot = plot +theme(strip.text = element_text(colour = 'white', face = "bold"))
plot = plot+ theme(legend.position="top",legend.box.spacing = margin(0.2))
  plot <- ggplot_gtable(ggplot_build(plot))
  strips <- which(grepl('strip-', plot$layout$name))
  pal <- viral_family_colors
for (i in seq_along(strips)) {
    k <- which(grepl('rect', plot$grobs[[strips[i]]]$grobs[[1]]$childrenOrder))
    l <- which(grepl('titleGrob', plot$grobs[[strips[i]]]$grobs[[1]]$childrenOrder))
    plot$grobs[[strips[i]]]$grobs[[1]]$children[[k]]$gp$fill <- pal[i]
}
plot(plot)

ggsave(plot, filename = paste0(fig3_path,"Fig_3A",".png"), height = 6, width = 10.5)
```


## saving virus x time splits
```{r}
dat_savePath2 = paste0(datPath,"generated_objs/")

virus = "arena"
arena_dat = list(arena_t1_dat, arena_t2_dat, arena_t3_dat, arena_t4_dat, arena_t5_dat)
names(arena_dat) = c("t1","t2","t3","t4","t5")
saveRDS(arena_dat, file = paste0(datPath,virus,"_timesplit_dat",".rds"))

virus = "corona"
corona_dat = list(corona_t1_dat, corona_t2_dat, corona_t3_dat, corona_t4_dat, corona_t5_dat)
names(corona_dat) = c("t1","t2","t3","t4","t5")
saveRDS(corona_dat, file = paste0(datPath,virus,"_timesplit_dat",".rds"))

virus = "filo"
filo_dat = list(filo_t1_dat, filo_t2_dat, filo_t3_dat, filo_t4_dat, filo_t5_dat)
names(filo_dat) = c("t1","t2","t3","t4","t5")
saveRDS(filo_dat, file = paste0(datPath,virus,"_timesplit_dat",".rds"))

virus = "flavi"
flavi_dat = list(flavi_t1_dat, flavi_t2_dat, flavi_t3_dat, flavi_t5_dat)
names(flavi_dat) = c("t1","t2","t3","t5")
saveRDS(flavi_dat, file = paste0(datPath,virus,"_timesplit_dat",".rds"))

virus = "ortho"
ortho_dat = list(ortho_t1_dat, ortho_t2_dat, ortho_t3_dat, ortho_t5_dat)
names(ortho_dat) = c("t1","t2","t3","t5")
saveRDS(ortho_dat, file = paste0(datPath,virus,"_timesplit_dat",".rds"))
```
