---
title: "01_3A"
output: html_document
---
---
title: "Generating Scores (Fig3)"
output: html_document
---

## script to plot each object by MVS score/ day and make table of mvs counts
```{r}
library(MetaIntegrator)
library(tidyr)
library(stringr)
library(dplyr)
library(tidyverse)
library(data.table)
library(ggpubr)
library(NatParksPalettes)
library(magrittr)

set.seed(4)
```

```{r setup_paths}
path = ""
datPath = paste0(path, "data/processed_data/")
datTable = paste0(path, "data/data_tables/")
source(paste0(path,"scripts/load_mvs_genes.R"))
source(paste0(path,"scripts/functions_step45.R"))
source(paste0(path,"scripts/subset_dataobjects.R"))
source(paste0(path,"scripts/run_pooled_aurocs.R"))
fig_path = paste0(path,"figures/")
fig3_path = paste0(fig_path, "Figure3/")
```

## starting object
# doing score generation on peak point for by DEGs from 3A
```{r}
dat = readRDS(paste0(datPath, "5a_scores_added.rds"))
time_to_keep = readRDS(paste0(datPath, "times_to_keep_2a.rds"))

time_to_keep$time_category = ifelse(time_to_keep$dataset5 %in% c("Coronaviridae","Orthomyxoviridae"), "t2","t4")
time_to_keep$time_category = ifelse(time_to_keep$dataset5 == "Flaviviridae", "t3", time_to_keep$time_category)
time_to_keep$time_category = ifelse(time_to_keep$obj_name == "GSE5790-LCMV_1", "t3", time_to_keep$time_category) ## there isn't animals at t4 but t3 only
time_to_keep$time_category = ifelse(time_to_keep$obj_name == "GSE24943-EBOV_2", "t3", time_to_keep$time_category)
time_to_keep$time_category = ifelse(time_to_keep$obj_name == "PRJNA718880-EBOV_3", "t3", time_to_keep$time_category)
time_to_keep$time_category = ifelse(time_to_keep$obj_name == "PRJNA398558-EBOV_4", "t3", time_to_keep$time_category)

dat = subset_data_with_dfkey(dat, time_to_keep)

saveRDS(dat, paste0(datPath, "16a_timecat_peaksubset_unbiased.rds"))
```

```{r}
all = readRDS(paste0(datPath, "16a_timecat_peaksubset_unbiased.rds"))

arena_dat = reformat(all, arena,  reMetanalaysis = TRUE)
arena_dat  = fix_formattedname(arena_dat)
corona_dat = reformat(all, corona,  reMetanalaysis = TRUE)
corona_dat  = fix_formattedname(corona_dat)
filo_dat = reformat(all, filo,  reMetanalaysis = TRUE)
filo_dat  = fix_formattedname(filo_dat)
flavi_dat = reformat(all, flavi,  reMetanalaysis = TRUE)
flavi_dat  = fix_formattedname(flavi_dat)
ortho_dat = reformat(all, ortho,  reMetanalaysis = TRUE)
ortho_dat  = fix_formattedname(ortho_dat)
all = readRDS(paste0(datPath, "16a_timecat_peaksubset_unbiased.rds")) %>% runMetaAnalysis()

virus_dat_list = list(arena_dat, corona_dat, filo_dat, flavi_dat,ortho_dat)
names(virus_dat_list) = c("arena", "corona", "filo", "flavi","ortho")
saveRDS(virus_dat_list, file = paste0(datPath, "11_peak_virus_metaanalysis_datlist.rds"))
saveRDS(all, file = paste0(datPath, "fig3_allviruses_scores_NHPpeak.rds"))
```

## generating scores using the objects created in 01_generating degs
```{r}
# arena = readRDS(file = paste0(datPath,"arena","_timesplit_dat.rds"))$t4 %>% runMetaAnalysis()
# corona = readRDS(file = paste0(datPath,"corona","_timesplit_dat.rds"))$t2 %>% runMetaAnalysis()
# filo = readRDS(file = paste0(datPath,"filo","_timesplit_dat.rds"))$t3 %>% runMetaAnalysis()
# flavi = readRDS(file = paste0(datPath,"flavi","_timesplit_dat.rds"))$t3 %>% runMetaAnalysis()
# ortho = readRDS(file = paste0(datPath,"ortho","_timesplit_dat.rds"))$t2 %>% runMetaAnalysis()
# all = readRDS(file = paste0(datPath, "16a_timecat_peaksubset_unbiased.rds")) %>% runMetaAnalysis()
# 
# 
# virus_dat_list = list(arena, corona, filo, flavi,ortho)
# names(virus_dat_list) = c("arena", "corona", "filo", "flavi","ortho")
# saveRDS(virus_dat_list, file = paste0(datPath, "11_peak_virus_metaanalysis_datlist.rds"))
# saveRDS(all, file = paste0(datPath, "fig3_allviruses_scores_NHPpeak.rds"))
```

## with effect size cutoffs
```{r}
run_filter_genes = function(obj, loos = c(TRUE,FALSE), fdr_thresh = c(.05,.01,.001, .0001), es_list = c(.6, 1)) {
  for (loo in loos){
    for (val in fdr_thresh){
      print(val)
      for (val2 in es_list){
        print(val2)
        obj = filterGenes(obj, isLeaveOneOut = loo, FDRThresh = val, effectSizeThresh = val2)
      }
    }
  }
  return(obj)
}

virus_dat_list = readRDS(file = paste0(datPath, "11_peak_virus_metaanalysis_datlist.rds"))
all = readRDS(paste0(datPath, "fig3_allviruses_scores_NHPpeak.rds"))

arena = virus_dat_list$arena %>% run_filter_genes()
corona = virus_dat_list$corona %>% run_filter_genes()
filo = virus_dat_list$filo %>% run_filter_genes()
flavi = virus_dat_list$flavi %>% run_filter_genes()
ortho = virus_dat_list$ortho %>% run_filter_genes()

all %<>% run_filter_genes()

virus_dat_list2 = list(arena, corona, filo, flavi, ortho,  all)
names(virus_dat_list2) = c("arena", "corona", "filo", "flavi", "ortho",  "all")

saveRDS(virus_dat_list2, file = paste0(datPath, "12a_scores_virus_datlist.rds"))
```


```{r}
virus_dat_list2 = readRDS(paste0(datPath, "12a_scores_virus_datlist.rds"))

### arena
cols = c("virus", "filter","pos","neg", "total")
arena_df <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(arena_df) <- cols

virus = "arena"
arena_obj = virus_dat_list2$arena

for (filter in names(arena_obj$filterResults)) {
  item = arena_obj$filterResults[[filter]]
  pos_num = length(item$posGeneNames)
  neg_num = length(item$negGeneNames)
  total_num = pos_num + neg_num
  row = c(virus,filter, pos_num,neg_num, total_num)
  arena_df = rbind(arena_df, row)
  colnames(arena_df) = cols
  arena_df$pos = as.numeric(arena_df$pos)
  arena_df$neg = as.numeric(arena_df$neg)
  arena_df$total = as.numeric(arena_df$total)
}

## corona
cols = c("virus", "filter","pos","neg", "total")
corona_df <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(corona_df) <- cols

virus = "corona"
corona_obj = virus_dat_list2$corona

for (filter in names(corona_obj$filterResults)) {
  item = corona_obj$filterResults[[filter]]
  pos_num = length(item$posGeneNames)
  neg_num = length(item$negGeneNames)
  total_num = pos_num + neg_num
  row = c(virus,filter, pos_num,neg_num, total_num)
  corona_df = rbind(corona_df, row)
  colnames(corona_df) = cols
  corona_df$pos = as.numeric(corona_df$pos)
  corona_df$neg = as.numeric(corona_df$neg)
  corona_df$total = as.numeric(corona_df$total)
}

## filo
cols = c("virus", "filter","pos","neg", "total")
filo_df <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(filo_df) <- cols

virus = "filo"
filo_obj = virus_dat_list2$filo

for (filter in names(filo_obj$filterResults)) {
  item = filo_obj$filterResults[[filter]]
  pos_num = length(item$posGeneNames)
  neg_num = length(item$negGeneNames)
  total_num = pos_num + neg_num
  row = c(virus,filter, pos_num,neg_num, total_num)
  filo_df = rbind(filo_df, row)
  colnames(filo_df) = cols
  filo_df$pos = as.numeric(filo_df$pos)
  filo_df$neg = as.numeric(filo_df$neg)
  filo_df$total = as.numeric(filo_df$total)
}

## flavi
cols = c("virus", "filter","pos","neg", "total")
flavi_df <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(flavi_df) <- cols

virus = "flavi"
flavi_obj = virus_dat_list2$flavi

for (filter in names(flavi_obj$filterResults)) {
  item = flavi_obj$filterResults[[filter]]
  pos_num = length(item$posGeneNames)
  neg_num = length(item$negGeneNames)
  total_num = pos_num + neg_num
  row = c(virus,filter, pos_num,neg_num, total_num)
  flavi_df = rbind(flavi_df, row)
  colnames(flavi_df) = cols
  flavi_df$pos = as.numeric(flavi_df$pos)
  flavi_df$neg = as.numeric(flavi_df$neg)
  flavi_df$total = as.numeric(flavi_df$total)
}

## ortho
cols = c("virus", "filter","pos","neg", "total")
ortho_df <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(ortho_df) <- cols

virus = "ortho"
ortho_obj = virus_dat_list2$ortho

for (filter in names(ortho_obj$filterResults)) {
  item = ortho_obj$filterResults[[filter]]
  pos_num = length(item$posGeneNames)
  neg_num = length(item$negGeneNames)
  total_num = pos_num + neg_num
  row = c(virus,filter, pos_num,neg_num, total_num)
  ortho_df = rbind(ortho_df, row)
  colnames(ortho_df) = cols
  ortho_df$pos = as.numeric(ortho_df$pos)
  ortho_df$neg = as.numeric(ortho_df$neg)
  ortho_df$total = as.numeric(ortho_df$total)
}


## all
cols = c("virus", "filter","pos","neg", "total")
all_df <- data.frame(matrix(ncol = 5, nrow = 0))
colnames(all_df) <- cols

virus = "all"
all_obj = virus_dat_list2$all

for (filter in names(all_obj$filterResults)) {
  item = all_obj$filterResults[[filter]]
  pos_num = length(item$posGeneNames)
  neg_num = length(item$negGeneNames)
  total_num = pos_num + neg_num
  row = c(virus,filter, pos_num,neg_num, total_num)
  all_df = rbind(all_df, row)
  colnames(all_df) = cols
  all_df$pos = as.numeric(all_df$pos)
  all_df$neg = as.numeric(all_df$neg)
  all_df$total = as.numeric(all_df$total)
}


full_table = rbind(arena_df, corona_df, filo_df, flavi_df, ortho_df, all_df)
full_table$pos = as.numeric(full_table$pos)
full_table$neg = as.numeric(full_table$neg)
full_table$total = as.numeric(full_table$total)
full_table2 = full_table
full_table2$dist = abs(200 - full_table2$total)
full_table2b = full_table2[full_table2$dist < 100,]
full_table2b = full_table2b[grepl(0.6,full_table2b$filter),]
```

```{r}
arena_threshold = "FDR1e-04_es0.6_nStudies1_looaFALSE_hetero0" #207
corona_threshold = "FDR0.001_es0.6_nStudies1_looaFALSE_hetero0" #181
filo_threshold = "FDR0.01_es0.6_nStudies1_looaTRUE_hetero0" #152
flavi_threshold = "FDR0.001_es0.6_nStudies1_looaFALSE_hetero0" #120
ortho_threshold = "FDR0.05_es0.6_nStudies1_looaFALSE_hetero0" #169
all_threshold = "FDR1e-04_es0.6_nStudies1_looaTRUE_hetero0" #257

score_thresholds = list(arena_threshold, corona_threshold, filo_threshold,
                        flavi_threshold, ortho_threshold,
                        all_threshold)
names(score_thresholds) = c("arena", "corona", "filo", "flavi", "ortho",  "all")

saveRDS(score_thresholds, file = paste0(datPath, "score_thresholds_string.rds"))
```


## pick the right score
```{r}
score_thresholds = readRDS(paste0(datPath, "score_thresholds_string.rds"))
virus_dat_list = readRDS(paste0(datPath, "12a_scores_virus_datlist.rds"))

arena_pos = virus_dat_list$arena$filterResults[[score_thresholds$arena]]$posGeneNames
arena_neg = virus_dat_list$arena$filterResults[[score_thresholds$arena]]$negGeneNames

corona_pos = virus_dat_list$corona$filterResults[[score_thresholds$corona]]$posGeneNames
corona_neg = virus_dat_list$corona$filterResults[[score_thresholds$corona]]$negGeneNames

filo_pos = virus_dat_list$filo$filterResults[[score_thresholds$filo]]$posGeneNames
filo_neg = virus_dat_list$filo$filterResults[[score_thresholds$filo]]$negGeneNames

flavi_pos = virus_dat_list$flavi$filterResults[[score_thresholds$flavi]]$posGeneNames
flavi_neg = virus_dat_list$flavi$filterResults[[score_thresholds$flavi]]$negGeneNames

ortho_pos = virus_dat_list$ortho$filterResults[[score_thresholds$ortho]]$posGeneNames
ortho_neg = virus_dat_list$ortho$filterResults[[score_thresholds$ortho]]$negGeneNames

all_pos = virus_dat_list$all$filterResults[[score_thresholds$all]]$posGeneNames
all_neg = virus_dat_list$all$filterResults[[score_thresholds$all]]$negGeneNames

scores_list = list(arena_pos, arena_neg, corona_pos, corona_neg, filo_pos, filo_neg,
                   flavi_pos, flavi_neg, ortho_pos, ortho_neg,
                   all_pos, all_neg,  mvs.up, mvs.down)
names(scores_list) = c("arena_pos", "arena_neg", "corona_pos", "corona_neg", "filo_pos", "filo_neg",
                   "flavi_pos", "flavi_neg", "ortho_pos", "ortho_neg",
                   "all_pos", "all_neg","MVS_pos","MVS_down")

saveRDS(scores_list, file = paste0(datPath, "score_genes.rds"))
```

# get table with effectsize, score, gene, direction
```{R}
scores_list = readRDS(paste0(datPath, "score_genes.rds"))
virus_dat_list2 = readRDS(paste0(datPath, "12a_scores_virus_datlist.rds"))

arena_df = virus_dat_list2$arena$metaAnalysis$pooledResults[c(scores_list$arena_pos,scores_list$arena_neg),]
arena_df$gene = rownames(arena_df)
arena_df = arena_df[,c("gene","effectSize")]
arena_df$score_type = "Arena Sig."

corona_df = virus_dat_list2$corona$metaAnalysis$pooledResults[c(scores_list$corona_pos,scores_list$corona_neg),]
corona_df$gene = rownames(corona_df)
corona_df = corona_df[,c("gene","effectSize")]
corona_df$score_type = "Corona Sig."

filo_df = virus_dat_list2$filo$metaAnalysis$pooledResults[c(scores_list$filo_pos,scores_list$filo_neg),]
filo_df$gene = rownames(filo_df)
filo_df = filo_df[,c("gene","effectSize")]
filo_df$score_type = "Filo Sig."

flavi_df = virus_dat_list2$flavi$metaAnalysis$pooledResults[c(scores_list$flavi_pos,scores_list$flavi_neg),]
flavi_df$gene = rownames(flavi_df)
flavi_df = flavi_df[,c("gene","effectSize")]
flavi_df$score_type = "Flavi Sig."

ortho_df = virus_dat_list2$ortho$metaAnalysis$pooledResults[c(scores_list$ortho_pos,scores_list$ortho_neg),]
ortho_df$gene = rownames(ortho_df)
ortho_df = ortho_df[,c("gene","effectSize")]
ortho_df$score_type = "Ortho Sig."

all_df = virus_dat_list2$all$metaAnalysis$pooledResults[c(scores_list$all_pos,scores_list$all_neg),]
all_df$gene = rownames(all_df)
all_df = all_df[,c("gene","effectSize")]
all_df$score_type = "VRS Sig."


signature_df = rbind(arena_df, corona_df, filo_df, flavi_df, ortho_df, all_df)
write.csv(signature_df, file = paste0(datTable,"stable5_scores.csv"), row.names=FALSE)
```

## generate all the scores to add to the full object (5a)
## subset full object by 3A peaks
```{r}
scores_list = readRDS(paste0(datPath, "score_genes.rds"))
dat = readRDS(paste0(datPath, "5a_scores_added.rds"))

dat = add_score_to_pheno(dat, "arena_score", scores_list$arena_pos, scores_list$arena_neg)
dat = add_score_to_pheno(dat, "corona_score", scores_list$corona_pos, scores_list$corona_neg)
dat = add_score_to_pheno(dat, "filo_score", scores_list$filo_pos, scores_list$filo_neg)
dat = add_score_to_pheno(dat, "flavi_score", scores_list$flavi_pos, scores_list$flavi_neg)
dat = add_score_to_pheno(dat, "ortho_score", scores_list$ortho_pos, scores_list$ortho_neg)
dat = add_score_to_pheno(dat, "all_score", scores_list$all_pos, scores_list$all_neg)

for (dataset in names(dat$originalData)) {
  if(unique(dat$originalData[[dataset]]$pheno$dataset5) == "Arenaviridae") {
    print("arena")
    dat$originalData[[dataset]]$pheno$own_score = dat$originalData[[dataset]]$pheno$arena_score
  }
  if(unique(dat$originalData[[dataset]]$pheno$dataset5) == "Coronaviridae") {
    print("corona")
    dat$originalData[[dataset]]$pheno$own_score = dat$originalData[[dataset]]$pheno$corona_score
  }
  if(unique(dat$originalData[[dataset]]$pheno$dataset5) == "Filoviridae") {
    print("filo")
    dat$originalData[[dataset]]$pheno$own_score = dat$originalData[[dataset]]$pheno$filo_score
  }
  if(unique(dat$originalData[[dataset]]$pheno$dataset5) == "Flaviviridae") {
    print("flavi")
    dat$originalData[[dataset]]$pheno$own_score = dat$originalData[[dataset]]$pheno$flavi_score
  }
  if(unique(dat$originalData[[dataset]]$pheno$dataset5) == "Orthomyxoviridae") {
    print("ortho")
    dat$originalData[[dataset]]$pheno$own_score = dat$originalData[[dataset]]$pheno$ortho_score
  }
}

saveRDS(dat, file = paste0(datPath, "5a_alldata_virus_scores_added.rds"))
```

```{r}
dat = readRDS(paste0(datPath, "5a_alldata_virus_scores_added.rds"))
time_to_keep = readRDS(paste0(datPath, "times_to_keep_2a.rds"))
dat = subset_data_with_dfkey(dat, time_to_keep)
saveRDS(dat, paste0(datPath, "16a_timecat_peaksubset_with_virus_scores.rds"))
```

```{r}
## pick the earliest day in the time category
subset_data_by_anitime = function(object, df, colname = "obj_name") {
  data_out = list()
  object = fix_time_cat(object)
  for (i in 1:length(names(object$originalData))) {
    dataset = object$originalData[[i]]
    pheno = dataset$pheno
    name = unname(unlist(unique(pheno[colname])))
    df_one = df[df$obj_name == name,]
    df_one$ani_day = paste0(df_one$animal,"_",df_one$day)
    pheno$ani_day = paste0(pheno$animal,"_",pheno$day)

    ## subset pheno by animal+time
    pheno_sub = pheno[pheno$ani_day %in% df_one$ani_day,]
    print(paste0(name,"_",nrow(pheno),"_",nrow(pheno_sub)))
    ## subset by dataset by pheno
    dataset <- subsetOriginalData(dataset, 
                                  keepMe= c(rownames(pheno_sub)))
    message(checkDataObject(dataset, "Dataset"))
    object$originalData[[i]] = dataset
    }
  return(object)
}
```

```{r}
dat = readRDS(paste0(datPath, "5a_alldata_virus_scores_added.rds"))
df_single = readRDS(paste0(datPath, "df_single.rds"))
dat = subset_data_by_anitime(dat, df_single)

for (i in c(1:length(names(dat$originalData)))) {
  print(unique(dat$originalData[[i]]$pheno$day))
  print(unique(dat$originalData[[i]]$pheno$time_cat))
}

saveRDS(dat, paste0(datPath, "16a_timecat_peaksubset_with_virus_scores_individual.rds"))

table(table(df_single$animal)) # confirm max 2 timepoints
```


